\documentclass[Chap1bmain.tex]{subfiles}
\begin{document}

\section{Other Types of Studies}
\citet{lewis} categorize method comparison studies into three
different types.  The key difference between the first two is
whether or not a `gold standard' method is used. In situations
where one instrument or method is known to be `accurate and
precise', it is considered as the`gold standard' \citep{lewis}. A
method that is not considered to be a gold standard is referred to
as an `approximate method'. In calibration studies they are
referred to a criterion methods and test methods respectively.


\textbf{1. Calibration problems}. The purpose is to establish a
relationship between methods, one of which is an approximate
method, the other a gold standard. The results of the approximate
method can be mapped to a known probability distribution of the
results of the gold standard \citep{lewis}. (In such studies, the
gold standard method and corresponding approximate method are
generally referred to a criterion method and test method
respectively.) \citet*{BA83} make clear that their methodology is
not intended for calibration problems.

\bigskip \textbf{2. Comparison problems}. When two approximate
methods, that use the same units of measurement, are to be
compared. This is the case which the Bland-Altman methodology is
specfically intended for, and therefore it is the most relevant of
the three.

\bigskip \textbf{3. Conversion problems}. When two approximate
methods, that use different units of measurement, are to be
compared. This situation would arise when the measurement methods
use 'different proxies', i.e different mechanisms of measurement.
\citet{lewis} deals specifically with this issue. In the context
of this study, it is the least relevant of the three.

\citet[p.47]{DunnSEME} cautions that`gold standards' should not be
assumed to be error free. `It is of necessity a subjective
decision when we come to decide that a particular method or
instrument can be treated as if it was a gold standard'. The
clinician gold standard , the sphygmomanometer, is used as an
example thereof.  The sphygmomanometer `leaves considerable room
for improvement' \citep{DunnSEME}. \citet{pizzi} similarly
addresses the issue of glod standards, `well-established gold
standard may itself be imprecise or even unreliable'.


The NIST F1 Caesium fountain atomic clock is considered to be the
gold standard when measuring time, and is the primary time and
frequency standard for the United States. The NIST F1 is accurate
to within one second per 60 million years \citep{NIST}.

Measurements of the interior of the human body are, by definition,
invasive medical procedures. The design of method must balance the
need for accuracy of measurement with the well-being of the
patient. This will inevitably lead to the measurement error as
described by \citet{DunnSEME}. The magnetic resonance angiogram,
used to measure internal anatomy,  is considered to the gold
standard for measuring aortic dissection. Medical test based upon
the angiogram is reported to have a false positive reporting rate
of 5\% and a false negative reporting rate of 8\%. This is
reported as sensitivity of 95\% and a specificity of 92\%
\citep{ACR}.

In literature they are, perhaps more accurately, referred to as
`fuzzy gold standards' \citep{phelps}. Consequently when one of the methods is
essentially a fuzzy gold standard, as opposed to a `true' gold
standard, the comparison of the criterion and test methods should
be consider in the context of a comparison study, as well as of a
calibration study.

\end{document}

