---
title: "Coillte Forest Automated Reporting Tool"
output:
  html_document:
    toc: true
    toc_depth: 2
   

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load libraries
library(dplyr)        # data wrangling
library(cartogram)    # for the cartogram
library(ggplot2)      # to realize the plots
library(broom)        # from geospatial format to data frame
library(tweenr)       # to create transition dataframe between 2 states
library(gganimate)    # To realize the animation
library(maptools)     # world boundaries coordinates
library(viridis)      # for a nice color palette
library(kableExtra)
library(mapproj)
library(gridExtra)
library(DT)
library(plotly)
```

## Plotly R Open Source Graphing Library

Plotly's R graphing library makes interactive, publication-quality graphs. Examples of how to make line plots, scatter plots, area charts, bar charts, error bars, box plots, histograms, heatmaps, subplots, multiple-axes, and 3D (WebGL based) charts.

## Example Data: USA Agricultural Exports 2011


```{r,echo=FALSE}
df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")

head(df)
```






```{r echo = FALSE}
library(plotly)
df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")
df$hover <- with(df, paste(state, '<br>', "Beef", beef, "Dairy", dairy, "<br>",
                           "Fruits", total.fruits, "Veggies", total.veggies,
                           "<br>", "Wheat", wheat, "Corn", corn))
# give state boundaries a white border
l <- list(color = toRGB("red"), width = 2)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('blue')
)
```

```{r echo = FALSE}
p <- plot_geo(df, locationmode = 'USA-states') %>%
  add_trace(
    z = ~total.exports, text = ~hover, locations = ~code,
    color = ~total.exports, colors = 'Purples'
  ) %>%
  colorbar(title = "Millions USD") %>%
  layout(
    title = '2011 US Agriculture Exports by State<br>(Hover for breakdown)',
    geo = g
  )

```

```{r,echo=FALSE}
p
```


## Displaying the Raw Data

The table below depicts the raw data for several variables. Other variables are omitted for sake of space.

```{r,echo=FALSE}

df1 <- df %>% dplyr::select(code:dairy)
datatable(df1, options = list(
  searching = FALSE,
  pageLength = 5,
  lengthMenu = c(5, 10, 15, 20)
))
```
