
\documentclass[12pt, a4paper]{report}
\usepackage{natbib}
\usepackage{vmargin}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{subfigure}
%\usepackage{amscd}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{amsthm}
%\usepackage[dvips]{graphicx}
\bibliographystyle{chicago}
\renewcommand{\baselinestretch}{1.8}

% left top textwidth textheight headheight % headsep footheight footskip
\setmargins{3.0cm}{2.5cm}{15.5 cm}{23.5cm}{0.5cm}{0cm}{1cm}{1cm}

\pagenumbering{arabic}


\begin{document}
\author{Kevin O'Brien}
\title{Bland Altman Methodologies}
\date{\today}
\maketitle

\tableofcontents \setcounter{tocdepth}{3}

\chapter{Bland and Altman's approach to MCS}
\section*{Overview}
\begin{itemize}
\item Difference Plot
\item Limits of Agreement
\item Further Approaches
\item Criticisms
\end{itemize}
%-------------------------------------------------------------%
\section{Difference Plot}
\citet{BA83} recommended the use of graphical techniques to assess agreement.
Principally their method is calculating, for each pair of corresponding two methods of measurement of some underlying quantity, with no replicate measurements, the difference and mean.

Differences are then plotted against the mean.
Hopkins argued that the bias in a subsequent Bland-Altman plot was
due, in part, to using least-squares regression at the calibration
phase.


\subsection{Model Specification}
The model underpinning the Bland-Altman approach can be presented as follows:

The case-wise means

The case-wise differences $d_i = x_i - y_i$
\[
\Sigma = \left( \begin{array}{cc}
                  \sigma^2_1 + \sigma^2_2 &  \frac{1}{2}(\sigma^2_1 - \sigma^2_2) \\
                   \frac{1}{2}(\sigma^2_1 - \sigma^2_2) &   \sigma^2_b + \frac{1}{4}(\sigma^2_1 + \sigma^2_2)
                \end{array}\right)
\]

\subsection{Pitman-Morgan Testing}
\citet{Pitman} and \citet{morgan} separately developed a test for the equality of two variances. Bland and Altman demonstrate how a regression based test is equivalent to the pitman-Morgan test.
\newpage
%-------------------------------------------------------------%
\section{Limits of Agreement}
\citet{BA86} introduces an enhancement on their previous work; Limits of Agreement.

%-------------------------------------------------------------%
\section{Further Approaches}
\citet{BA99} provides a regression-based approach for dealing with the scenario of non-constancy.


%-------------------------------------------------------------%
\section{Criticisms}
\citet{DunnSEME} criticises the over-reliance of analysts on the Bland-Altman methodology.


\addcontentsline{toc}{section}{Bibliography}

\bibliography{transferbib}
\end{document}

