Method comparison studies are performed in order to prove equivalence between two measurement methods or instruments. 
The identification of outliers is an important part of data analysis as outliers can indicate serious errors in the 
measurement process. Common outlier tests proposed in the literature require a homogeneous sample distribution and 
homoscedastic random error variances. However, datasets in method comparison studies usually do not meet these assumptions. 
To overcome this problem, different data transformation methods are proposed in the literature. However, they will only 
be applicable if the random errors can be described by simple additive or multiplicative models. 
In this work, a new outlier test based on robust linear regression is proposed which provides a general 
solution to the above problem. The LORELIA (LOcal RELIAbility) residual test is based on a local, robust 
residual variance estimator, given as a weighted sum of the observed residuals. Outlier limits are
estimated from the actual data situation without making assumptions on the underlying error variance model. 
The performance of the new test is demonstrated in examples and simulations.

Method comparison studies are performed in order to prove equivalence between two measurement methods or instruments. The identification of outliers is an important part of data analysis as outliers can indicate serious errors in the measurement process. Common outlier tests 
proposed in the literature require a homogeneous sample distribution and homoscedastic random error variances. However, datasets in method comparison studies usually do not meet these assumptions. To overcome this problem, different data transformation methods are proposed in the literature. However, they will only be applicable if the random errors can be described by simple additive or multiplicative models. In this work, a new outlier test based on robust linear regression is proposed which provides a general solution to the above problem. The LORELIA (LOcal RELIAbility) residual test is based on a local, robust residual variance estimator, given as a weighted sum of the observed residuals. Outlier limits are estimated from the actual data situation without making assumptions on the underlying error variance model. The performance of the new test is demonstrated in examples and simulations.
