
\documentclass[12pt, a4paper]{article}
\usepackage{natbib}
\usepackage{vmargin}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{subfigure}
%\usepackage{amscd}
\usepackage{amssymb}
\usepackage{subfigure}
\usepackage{amsbsy}
\usepackage{amsthm, amsmath}
%\usepackage[dvips]{graphicx}
\bibliographystyle{chicago}
\renewcommand{\baselinestretch}{1.4}

% left top textwidth textheight headheight % headsep footheight footskip
\setmargins{3.0cm}{2.5cm}{15.5 cm}{23.5cm}{0.5cm}{0cm}{1cm}{1cm}

\pagenumbering{arabic}


\begin{document}
\author{Kevin O'Brien}
\title{Mixed Models for Method Comparison Studies}
\tableofcontents
\section{Note on Roy's paper}
\begin{enumerate}


\item Basic model:
\begin{center}
$ \boldsymbol{y_{i}} = \boldsymbol{X_{i}\beta}
+ \boldsymbol{Z_{i}b_{i}} + \boldsymbol{\epsilon_{i}}, \qquad i=1,\dots,n$ \\
$\boldsymbol{Z_{i}} \sim \mathcal{N}(\boldsymbol{0,\Sigma}),\quad
\boldsymbol{\epsilon_{i}} \sim \mathcal{N}(\boldsymbol{0, \sigma^2
\boldsymbol{I} })$
\end{center}

Assumptions are made about homoskedasticity.

\item General model:
\begin{center}
$ \boldsymbol{y_{i}} = \boldsymbol{X_{i}\beta}
+ \boldsymbol{Z_{i}b_{i}} + \boldsymbol{\epsilon_{i}}, \qquad i=1,\dots,n$ \\
$\boldsymbol{Z_{i}} \sim \mathcal{N}(\boldsymbol{0,\Psi}),\quad
\boldsymbol{\epsilon_{i}} \sim \mathcal{N}(\boldsymbol{0,\sigma^2 \boldsymbol{\Lambda} })$
\end{center}

Assumptions about homoskedasticity are relaxed \cite[pg.202]{pb}.





\item $\sigma^2 \boldsymbol{\Lambda}$ is the general form for the VC structure for residuals.

\item The response vector $\boldsymbol{y}_{i}$ comprises the observations of
the subject, as measured by two methods, taking three measurements each.
Hence a $6 \times 1$ random vector corresponding to the $i$th subject.
\begin{equation}
\boldsymbol{y}_{i} = (y_{i}^{j1},y_{i}^{Jj2},y_{i}^{j3},y_{i}^{s1},y_{i}^{s2},y_{i}^{s3}) \prime
\end{equation}

\item The number of replicates is $p$. A subject will have up to
 $2p$ measurements, for the two instrument case, i.e. $Max(n_{i}) = 2p$.
(Let $k$ denote number of instruments, which is assumed to be $2$
unless stated otherwise.) For the blood pressure data $p=3$.

\item $\boldsymbol{\Psi}$ refers to the between-subject sources of variation. $\boldsymbol{R}_{i}$ refers to the within-subject
source of variation between two methods. LME models allow for the explicit analysis of each.

\item $\boldsymbol{\Psi}$ is the variance covariance structure for the random effects.


There is three alternative structures for
$\boldsymbol{\Psi}$, the diagonal form, the identity form and the general form.
\[
\boldsymbol{\Psi} =
\left(%
\begin{array}{c c}
  \psi^2_1 & 0  \\
  0 & \psi^2_2  \\
\end{array}%
\right)\qquad \mathrm{or} \qquad \boldsymbol{\Psi} =
\left(%
\begin{array}{c c}
  \psi_{11} & \psi_{12}  \\
  \psi_{21} & \psi_{22}  \\
\end{array}%
\right)
\qquad \mathrm{or} \qquad \boldsymbol{\Psi} =
\left(%
\begin{array}{c c}
  \psi_{11} & \psi_{12}  \\
  \psi_{21} & \psi_{22}  \\
\end{array}%
\right)
\]



\item $\boldsymbol{\epsilon}_{i}$ is a $n_{i}$-dimensional vector
comprised of residual components. For the blood pressure data $n_{i} = 85$.

\item $\boldsymbol{\beta}$ is the solutions of the means of the two methods. In the LME output, the bias ad corresponding
t-value and p-values are presented. This is relevant to Roy's first test.

\item $\boldsymbol{b}_{i}$ is a $m-$dimensional vector comprised of
the random effects.
\begin{equation}
\boldsymbol{b}_{i} = \left( \begin{array}{c}
  b_{1i} \\
  b_{21}  \\
\end{array}\right)
\end{equation}

\item $\boldsymbol{\Psi}$ is the variance-covariance matrix of the random effects ,
with $2 \times 2$ dimensions.
\begin{equation}
\boldsymbol{\Psi} =
\left(%
\begin{array}{c c}
  \psi_{11} & \psi_{12}  \\
  \psi_{21} & \psi_{22}  \\
\end{array}%
\right)
\end{equation}

\item  $\boldsymbol{\Sigma}$ represents the partial VC matrix of
the established matrix and the new method for any replicates.

\begin{equation}
\Sigma = \left( \begin{array}{cc}
  \sigma^2_{e} & \sigma^{en} \\
  \sigma_{en} & \sigma^2_{n} \\
\end{array}\right)
\end{equation}

\begin{itemize}
\item $\sigma^2_{e}$ - partial variance of the established method.
\item $\sigma^2_{n}$ - partial variance of the new method.
\item $\sigma_{en}$ - partial covariance between both methods.
\end{itemize}

\item $\boldsymbol{V}$ represents the correlation matrix of the replicated measurements on a given method.
$\boldsymbol{\Sigma}$ is the within-subject VC matrix.

\item $\boldsymbol{V}$ and $\boldsymbol{\Sigma}$ are positive
definite matrices. The dimensions of $\boldsymbol{V}$ and
$\boldsymbol{\Sigma}$ are $3 \times 3 ( = p \times p )$ and $ 2 \times
2 (= k \times k)$.

\item It is assumed that $\boldsymbol{V}$ is the same for both methods and $\boldsymbol{\Sigma}$ is
the same for all replications.

\item $\boldsymbol{V} \bigotimes \boldsymbol{\Sigma}$ creates a $ 6 \times 6 ( = kp \times
kp)$ matrix.
$\boldsymbol{R}_{i}$ is a sub-matrix of this.

\item The variance covariance structure $\boldsymbol{R}_{i}$ has a separable covariance structure.

\item The overall variability $\boldsymbol{\mbox{Block} \Omega_{i}}$ is the sum of the
between-subject variability $\boldsymbol{\Psi}$
and the within subject variability $\boldsymbol{\Sigma}$

\begin{equation}
\left( \begin{array}{cc}
  \omega^2_{e} & \omega^{en} \\
  \omega_{en} & \omega^2_{n} \\
\end{array}\right)
=
\left( \begin{array}{cc}
  \psi^2_{e} & \psi^{en} \\
  \psi_{en} & \psi^2_{n} \\
\end{array}\right)
+
\left( \begin{array}{cc}
  \sigma^2_{e} & \sigma^{en} \\
  \sigma_{en} & \sigma^2_{n} \\
\end{array}\right)
\end{equation}

\item  No special form of the random effects VC matrix $\boldsymbol{\Psi}$ is assumed.
Form cans be specified.
The \tt{pdMat} class is used by the `nlme' package to specify patterned VC matrices.
 \begin{itemize}
 \item pdDiag - Assumes random effects are independent, with different variance.
 \item pdIdent - Assumes random effects are independent, with same variance.
 \item pdSymm - General symmetric positive definite matrix.
 \item pdCompSymm
 \end{itemize}

\end{enumerate}

\newpage
\section{Lambda Structure}

\begin{equation}
\boldsymbol{\epsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma^2 \boldsymbol{\Lambda})
\end{equation}
\begin{enumerate}
\item A simple assumption is to assumes that residuals are independent and homoscedastic, i.e. $\boldsymbol{Lambda = I}$.

\item For the Bland Altman blood pressure data,$\boldsymbol{\Lambda}$ has kronecker product structure
and has dimensions $6 \times 6$.
\end{enumerate}




\section{Basic Models Fits}
Further to \citet{PB}, several simple LME models are constructed
for the blood pressure data. This data set is the subject of a
method comparison study in \citet{BA99}.

\subsection{Implementing the Mixed Models Fits}
They are implemented using the following {\tt{R}} code, utilising the
`nlme' package. An analysis of variance is used to compare the model fits.

The {\tt{R}} script:
\begin{verbatim}
fit1 = lme( BP ~ method, data = dat, random = ~1 | subject )
fit2 = update(fit1, random = ~1 | subject/method )
fit3 = update(fit1, random = ~method - 1 | subject )
#analysis of variance
anova(fit1,fit2,fit3)
\end{verbatim}


\begin{enumerate}


\item Simplest workable model, allows differences between methods
and incorporates a random intercept for each subject. For subject
1 we have
\[
\boldsymbol{X}_i =
\left(%
\begin{array}{cc}
  1 & 0 \\
  1 & 0 \\
  1 & 0 \\
  1 & 1 \\
  1 & 1 \\
  1 & 1 \\
\end{array}%
\right),\quad
\boldsymbol{\beta} =
\left(%
\begin{array}{c}
  \beta_0 \\
  \beta_1 \\
\end{array}%
\right), \quad
\boldsymbol{Z}_i =
\left(%
\begin{array}{c}
  1 \\
  1 \\
  1 \\
  1 \\
  1 \\
  1 \\
\end{array}%
\right), \quad \boldsymbol{b}_i = b
\]
where $\mathrm{E}(b)=0$ and $\mathrm{var}(b)=\psi.$

\item
\[
\boldsymbol{Z}_i =
\left(%
\begin{array}{c c}
  1 & 0 \\
  1 & 0 \\
  1 & 0 \\
  0 & 1 \\
  0 & 1 \\
  0 & 1 \\
\end{array}%
\right)
\quad \boldsymbol{b}_i =
\left(%
\begin{array}{c c}
  b_1 & 0  \\
  0 & b_2  \\
\end{array}%
\right)
\]

where $\mathrm{E}(b_i)=0$ and $\mathrm{var}(\boldsymbol{b})=
\boldsymbol{\Psi}$.

The variance of error terms is a $6 \times 6$ matrix.

\end{enumerate}


\newpage
















\newpage


\subsection{Laird Ware Formulation}
\begin{equation*}
\boldsymbol{y_{i}} = \boldsymbol{X_{i}\beta}  + \boldsymbol{Z_{i}b_{i}} + \boldsymbol{\epsilon_{i}}, \qquad i=1,\dots,85
\end{equation*}
\begin{eqnarray*}
\boldsymbol{Z_{i}} \sim \mathcal{N}(\boldsymbol{0,\Psi}),\qquad
\boldsymbol{\epsilon_{i}} \sim \mathcal{N}(\boldsymbol{0,\sigma^2\Lambda})
\end{eqnarray*}

\newpage
\subsection{Model Fit 1}

This is a simple model with no interactions. There is a fixed effect for each method and a random effect for each subject.
\begin{equation*}
y_{ijk} = \beta_{j}  + b_{i} + \epsilon_{ijk}, \qquad i=1,\dots,2, j=1,\dots,85, k=1,\dots,3
\end{equation*}

\begin{eqnarray*}
b_{i} \sim \mathcal{N}(0,\sigma^2_{b}), \qquad \epsilon_{i} \sim \mathcal{N}(0,\sigma^2)
\end{eqnarray*}

\begin{verbatim}
Linear mixed-effects model fit by REML
  Data: dat
  Log-restricted-likelihood: -2155.853
  Fixed: BP ~ method
(Intercept)     methodS
  127.40784    15.61961

Random effects:
 Formula: ~1 | subject
        (Intercept) Residual
StdDev:    29.39085 12.44454

Number of Observations: 510
Number of Groups: 85
\end{verbatim}

\newpage
\subsection{Model Fit 2}

This is a simple model, this time with an interaction effect.
There is a fixed effect for each method. This model has random effects at two levels $b_{i}$ for the subject, and
another, $b_{ij}$, for the respective method within each subject.
\begin{equation*}
y_{ijk} = \beta_{j}  + b_{i} + b_{ij} + \epsilon_{ijk}, \qquad i=1,\dots,2, j=1,\dots,85, k=1,\dots,3
\end{equation*}
\begin{eqnarray*}
b_{i} \sim \mathcal{N}(0,\sigma^2_{1}), \qquad b_{ij} \sim \mathcal{N}(0,\sigma^2_{2}), \qquad \epsilon_{i} \sim \mathcal{N}(0,\sigma^2)
\end{eqnarray*}

In this model, the random interaction terms all have the same variance $\sigma^2_{2}$. These terms are assumed to be independent of each other, even
within the same subject.

\begin{verbatim}
Linear mixed-effects model fit by REML
  Data: dat
  Log-restricted-likelihood: -2047.714
  Fixed: BP ~ method
(Intercept)     methodS
  127.40784    15.61961

Random effects:
 Formula: ~1 | subject
        (Intercept)
StdDev:    28.28452

 Formula: ~1 | method %in% subject
        (Intercept) Residual
StdDev:    12.61562 7.763666

Number of Observations: 510
Number of Groups:
            subject method %in% subject
                 85                 170
\end{verbatim}


\newpage
\subsection{Model Fit 3}

This model is a more general model, compared to 'model fit 2'. This model treats the random interactions for each subject as a vector and
allows the variance-covariance matrix for that vector to be estimated from the set of all positive-definite matrices.
$\boldsymbol{y_{i}}$ is the entire response vector for the $i$th subject.
$\boldsymbol{X_{i}}$ and $\boldsymbol{Z_{i}}$  are the fixed- and random-effects design matrices respectively.
\begin{equation*}
\boldsymbol{y_{i}} = \boldsymbol{X_{i}\beta}  + \boldsymbol{Z_{i}b_{i}} + \boldsymbol{\epsilon_{i}}, \qquad i=1,\dots,85
\end{equation*}
\begin{eqnarray*}
\boldsymbol{Z_{i}} \sim \mathcal{N}(\boldsymbol{0,\Psi}),\qquad
\boldsymbol{\epsilon_{i}} \sim \mathcal{N}(\boldsymbol{0,\sigma^2\Lambda})
\end{eqnarray*}

For the first subject the response vector, $\boldsymbol{y_{1}}$, is:
\begin{table}[ht]
\begin{center}
\begin{tabular}{rrllr}
  \hline
observation & BP & subject & method & replicate \\
  \hline
1 & 100.00 & 1 & J &   1 \\
  86 & 106.00 & 1 & J &   2 \\
  171 & 107.00 & 1 & J &   3 \\
  511 & 122.00 & 1 & S &   1 \\
  596 & 128.00 & 1 & S &   2 \\
  681 & 124.00 & 1 & S &   3 \\
   \hline
\end{tabular}
\end{center}
\end{table}
\newpage
The fixed effects design matrix $\boldsymbol{X_{i}}$ is given by:
\begin{table}[ht]
\begin{center}
\begin{tabular}{r|r}
  \hline
  (Intercept) & method S \\
  \hline
 1 & 0 \\
 1 & 0 \\
 1 & 0 \\
 1 & 1 \\
 1 & 1 \\
 1 & 1 \\
   \hline
\end{tabular}
\end{center}
\end{table}

The random effects design matrix $\boldsymbol{Z_{i}}$ is given by:
\begin{table}[ht]
\begin{center}
\begin{tabular}{r|r}
  \hline
 method J & method S \\
  \hline
 1 & 0 \\
 1 & 0 \\
 1 & 0 \\
 0 & 1 \\
 0 & 1 \\
 0 & 1 \\
   \hline
\end{tabular}
\end{center}
\end{table}
\newpage
The following output was obtained.
\begin{verbatim}
Linear mixed-effects model fit by REML
  Data: dat
  Log-restricted-likelihood: -2047.582
  Fixed: BP ~ method
(Intercept)     methodS
  127.40784    15.61961

Random effects:
 Formula: ~method - 1 | subject
 Structure: General positive-definite, Log-Cholesky parametrization
         StdDev    Corr
methodJ  30.455093 methdJ
methodS  31.477237 0.835
Residual  7.763666

Number of Observations: 510
Number of Groups: 85

\end{verbatim}
\newpage
\addcontentsline{toc}{section}{Bibliography}

\bibliography{transferbib}
\end{document}
