\documentclass[12pt, a4paper]{report}
\usepackage{epsfig}
\usepackage{subfigure}
%\usepackage{amscd}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{amsthm}
%\usepackage[dvips]{graphicx}
\usepackage{natbib}
\bibliographystyle{chicago}
\usepackage{vmargin}
% left top textwidth textheight headheight
% headsep footheight footskip
\setmargins{3.0cm}{2.5cm}{15.5 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.5}
\pagenumbering{arabic}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{ill}[theorem]{Example}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{axiom}{Axiom}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{notation}{Notation}
\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\renewcommand{\thenotation}{}
\renewcommand{\thetable}{\thesection.\arabic{table}}
\renewcommand{\thefigure}{\thesection.\arabic{figure}}
\title{Research notes: linear mixed effects models}
\author{ } \date{ }


\begin{document}
\author{Kevin O'Brien}
\title{Regression-based LoAs}

\tableofcontents \setcounter{tocdepth}{2}
%-------------------------------------------------

\chapter{BA99}

\section{Regression-based Limits of Agreement} Assuming that
there will be no curvature in the scatter-plot, the methodology
regresses the difference of methods ($d$) on the average of those
methods ($a$) with a simple intercept slope model; $\hat{d} =
b_{0}+ b_{1}a.$ Should the slope $b_{1}$ be found to be
negligible, $\hat{d}$ takes the value $\bar{d}$.

The next step to take in calculating the limits is also a
regression, this time of the residuals as a function of the scale
of the measurements, expressed by the averages $a_{i}$;
 $ \hat{R} = c_{0}+ c_{1}a_{i}$

With reference to absolute values following a half-normal
distribution with mean $\sigma\sqrt{\frac{2}{\pi}}$, \citet{BA99} formulate the regression based limits of agreement as
follows
 \begin{equation}
  \hat{d} \pm 1.96\sqrt{\frac{\pi}{2}}\hat{R} = \hat{d} \pm 2.46\hat{R}
 \end{equation}

 %------------------------------------------------%
\newpage

\chapter{BXC2010}

\section{1. Introduction}

\section{2. Model for LoA}

$95\%$ prediction interval

\[ \bar{D} \pm 1.96 \times s.d.(D_i) \sqrt{\frac{n+1}{n}} \]

The correct factor is $\sigma^2_1 + \sigma^2_2 \frac{n+1}{n}$

\section{3.Non constant difference}

3.1 Model

$D_i = (\alpha_1 - \alpha_2) + (\beta_1 + \beta_2)\mu_i +(e_{1i} + e_{2i})$

3.2 Regression of differences on averages

$\beta_{2|1} = \frac{1-b/2}{1+b/2} \geq 1-b $

$Y_{2|1} = -a +(1-b)y_1  \pm 2 \tau $

\section{4. Worked Examples}
4.1 Blood Glucose (Plasma and Capillary)

\begin{itemize}
\item 46 non diabetic obese people at 120 minutes after a 75g oral glucose challenge
\item $D = -2.24 + 0.33A$ with residual standard deviation of 1.08
\item Prediction  interval for the difference of sizes.
\item $Y_{C|P}  = 1.92 + 0.71Y_N \pm 1.86$ and $Y_{P|C}  =2.69 +1.40Y_N \pm 2.60$
\end{itemize}
4.2 Plasma volume (Nadler Hurley)

\section{5. Why is it wrong to use the regression of the differences on the averages.}

5.1 Substantially wrong

5.2 Statistically wrong
It is assumed that the averages are independent of the error terms.

\[
\frac{\sigma_1 - \sigma_2}
{\sigma_1 + \sigma_2}
=
\frac{\beta_1 - \beta_2}
{\beta_1 + \beta_2}
\] \therefore \[
\frac{\sigma_1 - \sigma_2}
{\sigma_1 + \sigma_2}
=
\frac{\beta_1 - \beta_2}
{\beta_1 + \beta_2}
\]
5.3 Why are the limits straight lines
The prediction limits are straight lines because the estimation variance $\sigma2,1$ and $\beta2,1$ is ignored.

5.4 What is the relation to Standard regression
The model (2) is not a standard model

Classical regression models are based on the conditional distribution of one method given another.
5.5 What is the relation to Deming Regression
Deming Regression does not solve the prediction problem
unless we are willing to assume a known value for the ratio of the variances.
In studies without replicates, there is no information about the variance ratio for the two methods.
6. How wrong is it to do it anyway?

\section{7. Discussion}

\section{8. Recommendation}
\addcontentsline{toc}{section}{Bibliography}
\bibliography{2012bib}
\end{document} 