\subsection{Introduction}
\begin{itemize}
\item Comparing two methods of measurement is normally done by computing limits of agreement (LoA), i.e. prediction limits for
a future difference between measurements with the two methods. When the difference is not constant it is not clear what
this means, since the difference between the methods depends on the average; hence, unlike the case where the difference is
constant, LoA cannot directly be translated into a prediction interval for a measurement by one method given that of another.
\item The main point in the paper by Bland and Altman [1] is however different from the outlook in this paper; Bland and Altman
mainly discuss whether two methods of measurement can be used interchangeably and how to assess this with the help of
proper statistical methods to derive LoA, i.e. prediction limits for differences between two methods.
This paper takes as starting point that the classical LoA can be converted to a prediction interval for one method given a
measurement by the other (details in the next section). This sort of relationship can be shown in a plot as a line with slope 1
and prediction limits as lines also with slope 1; applicable for the prediction both from method 1 to method 2 and vice versa. In
the case of non-constant difference it would be desirable to be able to produce a similar plot, usable both ways. Thus, the aim
of this paper is to produce a conversion from one method to another that also applies in the case where the difference between
methods is not constant.
\item In this paper, I set up a proper model for data for method comparison studies which in the case of constant difference between
methods leads to the classical LoA, and in the case of linear bias gives a simple formula for the prediction. The paper only
addresses the situation where only one measurement by each method is available, although replicate measurements by each
method are desirable whenever possible [2]. Moreover, the situation with non-constant variance over the range of measurements
is not covered either.
\end{itemize}
\end{document}
