\section{Assessment of Agreements in Linear and Generalized Linear Mixed Models}

% http://indigo.uic.edu/handle/10027/9520
\begin{itemize}
\item Study of measuring agreement is intend to evaluate whether the readings from one rater/ measurement 
agree with those from other raters/measurements. 
In this dissertation, we are going to present a general method to assess agreement for a large 
variety of data with repeated measurements using linear and generalized linear mixed models. 
\item In the first place, a set of agreement statistics, including mean square deviation, concordance 
correlation coefficient, precision and accuracy coefficients, is presented for evaluating the 
intra-, inter-, and total-rater agreement in the multiple-rater and multiple-replications cases. 
\item Secondly, likelihood-based approaches are developed to estimate all the agreement statistics. 
Asymptotic properties of these estimates are also discussed for different data structures. 
\item Furthermore, our method has the merit of handling missing values and covariates naturally, 
and a new set of restricted agreement statistics is proposed in order to capture the true random 
variations and between-instrument effects adjusted for the covariate effects. 

\item Simulations for both linear and generalized linear mixed models are conducted to show the accuracy and effectiveness 
of our approaches. In the end, two industry datasets are evaluated using our approach. 
\item One is the cardiac function measurements used to determine the agreement between impedance cardiography and radionuclide 
ventriculography estimates, and the other one is an antihypertensive patch dataset given by FDA for assessing 
individual bioequivalence.
\end{itemize}