---
title: "``addGraticule``"
author: "Kevin O'Brien"
date: "28 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
```


Add a Graticule on the map see https://github.com/turban/

### Description
Add a Graticule on the map see https://github.com/turban/Leaflet.Graticule

### Usage
<pre><code>
addGraticule(map, interval = 20, sphere = FALSE, style = list(color =
"#333", weight = 1), layerId = NULL, group = NULL,
options = pathOptions(pointerEvents = "none", clickable = FALSE))
</code></pre>

### Arguments

* map a map widget object
* interval The spacing in map units between horizontal and vertical lines.
* sphere boolean. Default FALSE
* style path options for the generated lines. See http://leafletjs.com/reference-1.3.1.html#path-option
* layerId the layer id
* group the name of the group this layer belongs to.
* options the path options for the graticule layer

### Examples

```{r cars}
leaf <- leaflet() %>%
addTiles() %>%
addGraticule()
leaf
```
