---
title: "addLayersControl Add UI controls to switch layers on and off"
author: "Kevin O'Brien"
date: "28 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
```

Uses Leaflet's built-in layers control feature to allow users to choose one of several base layers, and
to choose any number of overlay layers to view.

### Usage
<pre><code>
addLayersControl(map, baseGroups = character(0),
overlayGroups = character(0), position = c("topright", "bottomright",
"bottomleft", "topleft"), options = layersControlOptions(),
data = getMapData(map))
layersControlOptions(collapsed = TRUE, autoZIndex = TRUE, ...)
removeLayersControl(map)
</code></pre>

### Arguments

map the map to add the layers control to
baseGroups character vector where each element is the name of a group. The user will be
able to choose one base group (only) at a time. This is most commonly used for
mostly-opaque tile layers.
overlayGroups character vector where each element is the name of a group. The user can turn
each overlay group on or off independently.

position position of control: "topleft", "topright", "bottomleft", or "bottomright"
options a list of additional options, intended to be provided by a call to layersControlOptions
data the data object from which the argument values are derived; by default, it is the
data object provided to leaflet() initially, but can be overridden
collapsed if TRUE (the default), the layers control will be rendered as an icon that expands
when hovered over. Set to FALSE to have the layers control always appear in its
expanded state.
autoZIndex if TRUE, the control will automatically maintain the z-order of its various groups
as overlays are switched on and off.
... other options for layersControlOptions()


### Examples
```{r pressure, echo=FALSE}

leaflet() %>%
addTiles(group = "OpenStreetMap") %>%
addProviderTiles("Stamen.Toner", group = "Toner by Stamen") %>%
addMarkers(runif(20, -75, -74), runif(20, 41, 42), group = "Markers") %>%
addLayersControl(
baseGroups = c("OpenStreetMap", "Toner by Stamen"),
overlayGroups = c("Markers")
)
```
