---
title: "addMeasure Add a measure control to the map."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
```


### Usage
<pre><code>
addMeasure(map, position = "topright", primaryLengthUnit = "feet",
secondaryLengthUnit = NULL, primaryAreaUnit = "acres",
secondaryAreaUnit = NULL, activeColor = "#ABE67E",
completedColor = "#C8F2BE", popupOptions = list(className =
"leaflet-measure-resultpopup", autoPanPadding = c(10, 10)), captureZIndex = 10000, localization = "en", decPoint = ".",
thousandsSep = ",")
</code></pre>

### Arguments
* map a map widget object
* position standard Leaflet control position options.
* primaryLengthUnit, secondaryLengthUnit
* units used to display length results. secondaryLengthUnit is optional. Valid values are "feet", "meters", "miles", and "kilometers".
* primaryAreaUnit, secondaryAreaUnit units used to display area results. secondaryAreaUnit is optional. Valid values
are "acres", "hectares", "sqmeters", and "sqmiles".
* activeColor base color to use for map features rendered while actively performing a measurement.
* Value should be a color represented as a hexadecimal string.
* completedColor base color to use for features generated from a completed measurement. Value should be a color represented as a hexadecimal string.
* popupOptions list of options applied to the popup of the resulting measure feature. Properties may be any standard Leaflet popup options.
* captureZIndex Z-index of the marker used to capture measure clicks. Set this value higher than
the z-index of all other map layers to disable click events on other layers while a measurement is active.
* localization Locale to translate displayed text. Available locales include en (default), cn, de,
es, fr, it, nl, pt, pt_BR, pt_PT, ru, and tr
* decPoint Decimal point used when displaying measurements. If not specified, values are defined by the localization.
* thousandsSep Thousands separator used when displaying measurements. If not specified, values are defined by the localization.

#### Value
modified map
### Examples


```{r Examples}
leaf <- leaflet() %>%
addTiles() %>%
# central park
fitBounds( -73.9, 40.75, -73.95, 40.8 ) %>%
addMeasure()
leaf
# customizing
leaf %>% addMeasure(
position = "bottomleft",
primaryLengthUnit = "meters",
primaryAreaUnit = "sqmeters",
activeColor = "#3D535D",
completedColor = "#7D4479",
localization = "de"

)
```
