---
title: "``leaflet::addRasterImage``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
```


addRasterImage Add a raster image as a layer

#### Description
Create an image overlay from a RasterLayer object. This is only suitable for small to medium sized rasters, as the entire image will be embedded into the HTML page (or passed over the websocket in a Shiny context).

### Usage
<pre><code>
addRasterImage(map, x, colors = if (raster::is.factor(x)) "Set1" else
"Spectral", opacity = 1, attribution = NULL, layerId = NULL, group = NULL, project = TRUE, method = c("auto", "bilinear",
"ngb"), maxBytes = 4 * 1024 * 1024, data = getMapData(map)) 

projectRasterForLeaflet(x, method)
</code></pre>

### Arguments

* ``map``: a map widget object
* ``x``: a RasterLayer object–see raster
* ``colors``: the color palette (see colorNumeric) or function to use to color the raster values
(hint: if providing a function, set ``na.color`` to "#00000000" to make NA areas transparent)
* ``opacity``: the base opacity of the raster, expressed from 0 to 1
* ``attribution``: the HTML string to show as the attribution for this layer
* ``layerId``: the layer id
* ``group``: the name of the group this raster image should belong to (see the same parameter under addTiles)
* ``project``: if TRUE, automatically project x to the map projection expected by Leaflet (EPSG:3857);
if FALSE, it’s the caller’s responsibility to ensure that x is already projected, and
that extent(x) is expressed in WGS84 latitude/longitude coordinates
* ``method``: the method used for computing values of the new, projected raster image. "bilinear" (the default) is appropriate for continuous data, "ngb" - nearest neighbor - is appropriate
for categorical data. Ignored if ``project = FALSE``. See projectRaster
for details.
* ``maxBytes``: the maximum number of bytes to allow for the projected image (before base64 encoding); defaults to 4MB.
* ``data``: the data object from which the argument values are derived; by default, it is the data object provided to ``leaflet()`` initially, but can be overridden

#### Details

The maxBytes parameter serves to prevent you from accidentally embedding an excessively large amount of data into your htmlwidget. This value is compared to the size of the final compressed image (after the raster has been projected, colored, and PNG encoded, but before base64 encoding is applied). Set ``maxBytes`` to ``Inf`` to disable this check, but be aware that very large rasters may not only make your map a large download but also may cause the browser to become slow or unresponsive.

By default, the addRasterImage function will project the RasterLayer x to EPSG:3857 using the raster package’s projectRaster function. This can be a time-consuming operation for even moderately sized rasters. Upgrading the raster package to 2.4 or later will provide a large speedup versus previous versions. If you are repeatedly adding a particular raster to your Leaflet maps, you can perform the projection ahead of time using ``projectRasterForLeaflet()``, and call ``addRasterImage`` with ``project = FALSE``.

#### Examples

```{r}
library(raster)
r <- raster(xmn = -2.8, xmx = -2.79, ymn = 54.04, ymx = 54.05, nrows = 30, ncols = 30)
values(r) <- matrix(1:900, nrow(r), ncol(r), byrow = TRUE)
crs(r) <- CRS("+init=epsg:4326")

if (requireNamespace("rgdal")) {
leaflet() %>% addTiles() %>%
  addRasterImage(r, colors = "Spectral", opacity = 0.8)
}
```
