---
title: "``ablakey/Leaflet.SimpleGraticule``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
```



### Description

Add a simple Graticule on the map 
* see https://github.com/ablakey/Leaflet.SimpleGraticule

### Usage

<pre><code>
addSimpleGraticule(map, interval = 20, showOriginLabel = TRUE,
redraw = "move", hidden = FALSE, zoomIntervals = list(),
layerId = NULL, group = NULL)
</code></pre>

### Arguments

* ``map``: a map widget object
* ``interval``: The spacing in map units between horizontal and vertical lines.
* ``showOriginLabel``: true Whether or not to show ’(0,0)’ at the origin.
* ``redraw``: on which map event to redraw the graticule. On move is default but moveend can be smoother.
* ``hidden``: hide on start
* ``zoomIntervals``: use different intervals in different zoom levels. If not specified, all zoom levels use value in interval option.
* ``layerId``: the layer id
* ``group``: the name of the group this layer belongs to.

### Examples

```{r}
leaflet() %>%
addTiles() %>%
addSimpleGraticule()
```
