\section{Probability-based Measures of Agreement}
There are two measures of agreement based on the probability criteria. The first is the
$p_0$-th percentile of jDj, say Q($p_0$), where $p_0$ (> 0.5) is a specified large probability (usually
¸ 0:80). 
It was introduced by Lin (2000) who called it the total deviation index (TDI). Its
small value indicates a good agreement between (X; Y ). The TDI can be expressed as,

! EQUATION HERE
-distribution with a single degree of freedom
and non-centrality parameter $\Del$.

%-----------------------------------------------------------------------------------------%
%Coverarge Probability

The second measure, introduced by Lin et al. (2002), is the \textbf{coverage probability} (CP) of
the interval [¡±0; ±0], where a difference under §±0 is considered practically equivalent to
zero. There is no loss of generality in taking this interval to be symmetric around zero as it
can be achieved by a location shift. 

Letting,
dl = (¡±0 ¡ ¹)=¾; du = (±0 ¡ ¹)=¾; (2)


the CP can be expressed as
F(±0) = ©(du) ¡ ©(dl): (3)

A high value of F(±0) implies a good agreement between the methods.


%-----------------------------------------------------------------------------------------%
\section{Coverage Probability and Tolerance Deviation Index}

Individual agreement between two measurement methods may be
assessed using the the coverage probability (CP) criteria or the
total deviation index (TDI) as proposed by \citet{lin2000} and
\citet{lin2002}.

If $d_{0}$ is predetermined as the maximum acceptable absolute
difference between two methods of measurement, the probability
that the absolute difference of two measures being less than
$d_{0}$ can be computed. This is known as the coverage probability
(CP).

\begin{equation}
CP = P(|x_{i} - y_{i}| \leq d_{0})
\end{equation}

If $\pi_{0}$ is set as the predetermined coverage probability, the
boundary under which the proportion of absolute differences is
$\pi_{0}$ may be determined. his boundary is known as the `total
deviation index' (TDI). Hence the TDI is the $100\pi_{0}$
percentile of the absolute difference of paired observations.


The CP is the most intuitively clear approach; it mirrors the information provided by the TDI. 
Both TDI and CP depend on the normality assumption and offer better power
for inference than the CCC. The CP would have difŽficulty discriminating among instruments or 
assays that have excellent agreement, all because the CP values would be very close to
1. In this case, the TDI can be used to discriminate among these. When a meaningful clinical range is known and the study is conducted over that range, the CCC offers a meaningful geo- metric interpretation and is unit free. Furthermore, the accuracy and precision components of the CCC offer more insight. Therefore, the CCC, accuracy, and precision remain very useful tools. Note that when Y and X are not linearly related, the CCC will capture the total deviation. However, it will treat the nonlinear deviation as imprecision rather than inaccuracy. The CCC, ICC, and Pearson correlation coefŽ cient depend
largely on the analytical range and the intrasample variation.

%-------------------------------------------------------------------------------------------%
\section{Mean Square Deviation}
Mean Square deviation is defined as the expectation of the squared difference of two readings.
The MSD is usually used for the case of two methods, each making a single reading.





