Maps with ggplot2
========================================================

### Loading Up The Relevant Packages

```{r}
library(maps)
library(ggplot2)
library(RColorBrewer)

states_map <- map_data("state")
```

Plot 1 : USA states Map

```{r fig.width=7, fig.height=6}
ggplot(states_map, 
       aes(x=long,y=lat,group=group)) + 
  geom_polygon(fill="white",colour="black")

```

Plot 2 : Projection

```{r fig.width=7, fig.height=6}

ggplot(states_map, aes(x=long,y=lat,group=group)) + 
  geom_path() + coord_map("mercator")

```

Plot 3 : Map of the World

You can also embed plots, for example:

```{r}

world_map <- map_data("world")
head(world_map)

```



You can also embed plots, for example:

```{r fig.width=7, fig.height=6}

east_asia <- map_data("world",region=c("Japan","China","North Korea","South Korea"))
ggplot(east_asia, aes(x=long,y=lat,group=group,fill=region)) + 
  geom_polygon(colours="black")+
  scale_fill_brewer(palette="Set2")
```

You can also embed plots, for example:

```{r fig.width=7, fig.height=6}
#Other Palettes : Accent,Paired, Set1


ggplot(east_asia, aes(x=long,y=lat,
                      group=group,fill=region)) + 
  geom_polygon(colours="black")+
  scale_fill_brewer(palette="Set1")
```
### Ireland
The main issue with this map is that it is constructed using a very small data set.

```{r}
IrlSet <- grep("Irel",world_map$region)
# IrlMap <- world_map[IrlSet,]
Ireland <- map_data("world",region=c("Ireland"))
```

```{r fig.width=7, fig.height=6}

ggplot(Ireland , aes(x=long,y=lat,group=group,fill=region)) + 
  geom_polygon(colours="black")+
  scale_fill_brewer(palette="Set2")
# Not Great !
```

### Some European Countries
The main issue with this map is that it is constructed using a very small data set.

```{r}

NWE <- map_data("world",region=c("Portugal","Ireland","Italy","Greece","Spain"))

```

```{r fig.width=7, fig.height=6}
ggplot(NWE , aes(x=long,y=lat,group=group,fill=region)) + 
  geom_polygon(colours="black")+
  scale_fill_brewer(palette="Set1")
```

```{r}

NZ1 <- map_data("world",region="New Zealand")
NZ1 <- subset(NZ1, long>0 &lat> -48)  #trim off outlying islands
```
```{r fig.width=7, fig.height=6}

ggplot(NZ1,aes(x=long,y=lat,group=group)) + geom_path()
```

Same map from a more detailed data source

```{r fig.width=7, fig.height=6}
NZ2 <- map_data("nz")
ggplot(NZ2,aes(x=long,y=lat,group=group)) + geom_path()

```

