Maps with ggplot2 : Choropleth Maps
========================================================

First, make sure packages are installed and loaded.

```{r}
#install.packages("plyr")
library(ggplot2)
library(maps)
library(plyr)
```

Set up our Data Set. Comprised of the the USA arrests data set and the states data.


```{r}
crimes <-data.frame(state =tolower(rownames(USArrests)),USArrests)
head(crimes)
```



```{r}
#library(maps)
states_map <- map_data("state")
crime_map <- merge(states_map,crimes,by.x="region",by.y="state")
head(crime_map)

```

### Crime Data Set
Let's look at the Assault data set in particular

```{r} 
crimes <-data.frame(state =tolower(rownames(USArrests)),USArrests)
head(crimes)
```

### Crime Data Set
Let's look at the Assault data set in particular

```{r} 

#library(maps)
states_map <- map_data("state")
crime_map <- merge(states_map,crimes,by.x="region",by.y="state")
head(crime_map)
# Use this package for the arrange function().
crime_map <- arrange(crime_map,group,order)
head(crime_map)
```

### Crime Data Set
Let's look at the Assault data set in particular

```{r} 
dim(crime_map)
summary(crime_map$Assault)
unique(crime_map$Assault)
#some states may have some value.
```





Choropleth Maps

```{r fig.width=7, fig.height=6}     
ggplot(crimes,aes(map_id=state,fill=Assault))+
   geom_map(map=states_map,colour="black")+ 
   scale_fill_gradient2(low="#559999",mid="grey90",high="#BB650B",
                            midpoint=median(crimes$Assault)) +
   expand_limits(x=states_map$long,y=states_map$lat) +
         coord_map("polyconic")
```

Choropleth Maps

```{r fig.width=7, fig.height=6}     
ggplot(crimes,aes(map_id=state,fill=Assault))+
geom_map(map=states_map,colour="black")+ 
scale_fill_gradient2(low="#559999",mid="grey90",high="#BB650B",
                              midpoint=median(crimes$Assault)) +
         expand_limits(x=states_map$long,y=states_map$lat) +
         coord_map("polyconic")
       

```