---
title: "``previewColors``: Color previewing utility"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
```


colorNumeric Color mapping

#### Description

Conveniently maps data values (numeric or factor/character) to colors according to a given palette,
which can be provided in a variety of formats.

### Usage

<pre><code>
colorNumeric(palette, domain, na.color = "#808080", alpha = FALSE,
reverse = FALSE)
colorBin(palette, domain, bins = 7, pretty = TRUE,
na.color = "#808080", alpha = FALSE, reverse = FALSE,
right = FALSE)
colorQuantile(palette, domain, n = 4, probs = seq(0, 1, length.out = n
+ 1), na.color = "#808080", alpha = FALSE, reverse = FALSE,
right = FALSE)
colorFactor(palette, domain, levels = NULL, ordered = FALSE,
na.color = "#808080", alpha = FALSE, reverse = FALSE)
</code></pre>

### Arguments
* palette The colors or color function that values will be mapped to
* domain The possible values that can be mapped. For colorNumeric and colorBin, this can be a simple numeric range (e.g. c(0, 100)); colorQuantile needs representative numeric data; and colorFactor
needs categorical data.
If NULL, then whenever the resulting color function is called, the x value will
represent the domain. This implies that if the function is invoked multiple times,the encoding between values and colors may not be consistent; if consistency is needed, you must provide a non-NULL domain.
* na.color The color to return for NA values. Note that na.color = NA is valid. alpha Whether alpha channels should be respected or ignored. If TRUE then colors without explicit alpha information will be treated as fully opaque.
* reverse Whether the colors (or color function) in palette should be used in reverseorder. For example, if the default order of a palette goes from blue to green, then reverse = TRUE will result in the colors going from green to blue.
* bins Either a numeric vector of two or more unique cut points or a single number (greater than or equal to 2) giving the number of intervals into which the domain values are to be cut.

* pretty Whether to use the function pretty() to generate the bins when the argument bins is a single number. When pretty = TRUE, the actual number of bins may not be the number of bins you specified. When pretty = FALSE, seq() is used to generate the bins and the breaks may not be "pretty".
* right parameter supplied to cut. See Details
* n Number of equal-size quantiles desired. For more precise control, use the probs argument instead.
* probs See quantile. If provided, the n argument is ignored.
* levels An alternate way of specifying levels; if specified, domain is ignored ordered If TRUE and domain needs to be coerced to a factor, treat it as already in the correct order

#### Details

colorNumeric is a simple linear mapping from continuous numeric data to an interpolated palette.
colorBin also maps continuous numeric data, but performs binning based on value (see the cut function). colorBin defaults for the cut function are include.lowest = TRUE and right = FALSE.
colorQuantile similarly bins numeric data, but via the quantile function.
colorFactor maps factors to colors. If the palette is discrete and has a different number of colors
than the number of factors, interpolation is used.

The palette argument can be any of the following:
1. A character vector of RGB or named colors. ### Examples: palette(), c("#000000", "#0000FF", "#FFFFFF"),
topo.colors(10)
2. The name of an RColorBrewer palette, e.g. "BuPu" or "Greens".
3. The full name of a viridis palette: "viridis", "magma", "inferno", or "plasma".
4. A function that receives a single value between 0 and 1 and returns a color. ### Examples:
colorRamp(c("#000000", "#FFFFFF"), interpolate = "spline").

#### Value
A function that takes a single parameter x; when called with a vector of numbers (except for
colorFactor, which expects factors/characters), #RRGGBB color strings are returned (unless
alpha = TRUE in which case #RRGGBBAA may also be possible).

### Examples

```{r}
pal <- colorBin("Greens", domain = 0:100)
pal(runif(10, 60, 100))
# Exponential distribution, mapped continuously
previewColors(colorNumeric("Blues", domain = NULL), sort(rexp(16)))
# Exponential distribution, mapped by interval
previewColors(colorBin("Blues", domain = NULL, bins = 4), sort(rexp(16)))
# Exponential distribution, mapped by quantile
previewColors(colorQuantile("Blues", domain = NULL), sort(rexp(16)))
createLeafletMap 27
# Categorical data; by default, the values being colored span the gamut...
previewColors(colorFactor("RdYlBu", domain = NULL), LETTERS[1:5])
# ...unless the data is a factor, without droplevels...
previewColors(colorFactor("RdYlBu", domain = NULL), factor(LETTERS[1:5], levels = LETTERS))
# ...or the domain is stated explicitly.
previewColors(colorFactor("RdYlBu", levels = LETTERS), LETTERS[1:5])
```
