---
title: "``leaflet::groupOptions ``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
```

#### Description

Change options on layer groups. Currently the only option is to control what zoom levels a layer group will be displayed at. The zoomLevels option is not compatible with layers control; do not both assign a group to zoom levels and use it with ``addLayersControl``.

#### Usage

<pre><code>

groupOptions(map, group, zoomLevels = NULL)

<code></pre>

#### Arguments

* ``map``: the map to modify

* ``group``: character vector of one or more group names to set options on

* ``zoomLevels``: numeric vector of zoom levels at which group(s) should be visible, or TRUE to display at all zoom levels

#### Examples

```{r}

pal <- colorQuantile("YlOrRd", quakes$mag)

leaflet() %>%
  addTiles(group = "basic") %>%    # Basic markers
  addMarkers(data = quakes, group = "basic") %>%
  addProviderTiles(providers$Stamen.TonerLite, group = "detail") %>%
  addCircleMarkers(data = quakes, group = "detail", fillOpacity = 0.5,
                   radius = ~mag * 5, color = ~pal(mag), stroke = FALSE) %>%
  groupOptions("detail", zoomLevels = 7:18)
```

When zoomed in, we'll show circles at the base of each marker whose radius and color reflect the magnitude
Also, we set the detail group to only appear when zoomed in


