---
title: "``leaflet::icons``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
```

## ``icons`` Create a list of icon data

#### Description
An icon can be represented as a list of the form list(iconUrl,iconSize, ...). This function is vectorized over its arguments to create a list of icon data. Shorter argument values will be re-cycled. NULL values for these arguments will be ignored.

### Usage

<pre><code>
icons(iconUrl = NULL, iconRetinaUrl = NULL, iconWidth = NULL,
iconHeight = NULL, iconAnchorX = NULL, iconAnchorY = NULL,
shadowUrl = NULL, shadowRetinaUrl = NULL, shadowWidth = NULL,
shadowHeight = NULL, shadowAnchorX = NULL, shadowAnchorY = NULL,
popupAnchorX = NULL, popupAnchorY = NULL, className = NULL)
</code></pre> 


#### Arguments

* ``iconUrl``: the URL or file path to the icon image
* ``iconRetinaUrl``: the URL or file path to a retina sized version of the icon image
* ``iconWidth``, ``iconHeight`` : size of the icon image in pixels
* ``iconAnchorX``, ``iconAnchorY`` : the coordinates of the "tip" of the icon (relative to its top left corner, i.e. the top
left corner means ``iconAnchorX = 0`` and ``iconAnchorY = 0``), and the icon will be aligned so that this point is at the marker’s geographical location
* ``shadowUrl``: the URL or file path to the icon shadow image
* ``shadowRetinaUrl``: the URL or file path to the retina sized version of the icon shadow image
* ``shadowWidth``, ``shadowHeight``: size of the shadow image in pixels
* ``shadowAnchorX``, ``shadowAnchorY``: the coordinates of the "tip" of the shadow
* ``popupAnchorX``, ``popupAnchorY``: the coordinates of the point from which popups will "open", relative to the icon anchor
* ``className``: a custom class name to assign to both icon and shadow images

### Value
A list of icon data that can be passed to the icon argument of addMarkers().

### Examples

```{r}
library(leaflet)
# adapted from http://leafletjs.com/### Examples/custom-icons.html
iconData <- data.frame(
lat = c(rnorm(10, 0), rnorm(10, 1), rnorm(10, 2)),
lng = c(rnorm(10, 0), rnorm(10, 3), rnorm(10, 6)),
group = rep(sort(c("green", "red", "orange")), each = 10),
stringsAsFactors = FALSE
)
```
```{r}
leaflet() %>% addMarkers(
  data = iconData,
  icon = ~ icons(
    iconUrl = sprintf("http://leafletjs.com/### Examples/custom-icons/leaf-%s.png", group),
    shadowUrl = "http://leafletjs.com/### Examples/custom-icons/leaf-shadow.png",
    iconWidth = 38, iconHeight = 95, shadowWidth = 50, shadowHeight = 64,
    iconAnchorX = 22, iconAnchorY = 94, shadowAnchorX = 4, shadowAnchorY = 62,
    popupAnchorX = -3, popupAnchorY = -76 
  )
)

# use point symbols from base R graphics as icons
pchIcons <- function(pch = 0:14, width = 30, height = 30, ...) {
n <- length(pch)
files <- character(n)
# create a sequence of png images
for (i in seq_len(n)) {
f <- tempfile(fileext = ".png")
png(f, width = width, height = height, bg = "transparent")
par(mar = c(0, 0, 0, 0))
plot.new()
points(.5, .5, pch = pch[i], cex = min(width, height) / 8, ...)
dev.off()
files[i] <- f
}
files

}
iconData <- matrix(rnorm(500), ncol = 2)
res <- kmeans(iconData, 10)
iconData <- cbind(iconData, res$cluster)
colnames(iconData) <- c("lat", "lng", "group")
iconData <- as.data.frame(iconData)
# 10 random point shapes for the 10 clusters in iconData
shapes <- sample(0:14, 10)
iconFiles <- pchIcons(shapes, 40, 40, col = "steelblue", lwd = 2)
# note the data has 250 rows, and there are 10 icons in iconFiles; they are
# connected by the `group` variable: the i-th row of iconData uses the
# group[i]-th icon in the icon list
```
```{r}
leaflet() %>% addMarkers(
data = iconData,
icon = ~ icons(
iconUrl = iconFiles[group],
popupAnchorX = 20, popupAnchorY = 0
),
popup = ~ sprintf(
"lat = %.4f, long = %.4f, group = %s, pch = %s", lat, lng, group, shapes[group]
)
)
unlink(iconFiles) # clean up the tmp png files that have been embedded
```

