---
title: "leafletOutput: Wrapper functions for using leaflet in shiny"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
```

#### Description
Use leafletOutput() to create a UI element, and renderLeaflet() to render the map widget.

### Usage
<pre><code>
leafletOutput(outputId, width = "100%", height = 400)
renderLeaflet(expr, env = parent.frame(), quoted = FALSE)
</code></pre> 

#### Arguments

* outputId output variable to read from width, height the width and height of the map (see shinyWidgetOutput)
* expr An expression that generates an HTML widget
* env The environment in which to evaluate expr.
* quoted Is expr a quoted expression (with quote())? This is useful if you want to save an expression in a variable.

### Examples
```{r}
# !formatR
library(shiny)
 app <- shinyApp(
    ui = fluidPage(leafletOutput('myMap')),
    server = function(input, output) {
      map = leaflet() %>% addTiles() %>% setView(-93.65, 42.0285, zoom = 17)
      output$myMap = renderLeaflet(map)
    }
 )
if (interactive()) app
```
