
---
title: "``leaflet-MiniMap``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
```

Add a minimap to the Map https://github.com/Norkart/Leaflet-MiniMap

### Usage

<pre><code>
addMiniMap(map, position = "bottomright", width = 150, height = 150,
  collapsedWidth = 19, collapsedHeight = 19, zoomLevelOffset = -5,
  zoomLevelFixed = FALSE, centerFixed = FALSE, zoomAnimation = FALSE,
  toggleDisplay = FALSE, autoToggleDisplay = FALSE,
  minimized = FALSE, aimingRectOptions = list(color = "#ff7800", weight= 1, clickable = FALSE), shadowRectOptions = list(color = "#000000",
  weight = 1, clickable = FALSE, opacity = 0, fillOpacity = 0),
  strings = list(hideText = "Hide MiniMap", showText = "Show MiniMap"),
  tiles = NULL, mapOptions = list())
</code></pre>

### Arguments
* map a map widget object
* position The standard Leaflet.Control position parameter, used like all the other controls.
Defaults to "bottomright".
* width The width of the minimap in pixels. Defaults to 150.
* height The height of the minimap in pixels. Defaults to 150.
* collapsedWidth The width of the toggle marker and the minimap when collapsed, in pixels.
Defaults to 19.
* collapsedHeight The height of the toggle marker and the minimap when collapsed, in pixels.
Defaults to 19.
* zoomLevelOffset The offset applied to the zoom in the minimap compared to the zoom of the
main map. Can be positive or negative, defaults to -5.
* ``zoomLevelFixed``: Overrides the offset to apply a fixed zoom level to the minimap regardless of
the main map zoom. Set it to any valid zoom level, if unset ``zoomLevelOffset`` is used instead.
* ``centerFixed``: Applies a fixed position to the minimap regardless of the main map’s view /
position. Prevents panning the minimap, but does allow zooming (both in the minimap and the main map). If the minimap is zoomed, it will always zoom around the centerFixed point. You can pass in a LatLng-equivalent object. Defaults to false.
* ``zoomAnimation``: Sets whether the minimap should have an animated zoom. (Will cause it to lag a bit after the movement of the main map.) Defaults to false.
* toggleDisplay Sets whether the minimap should have a button to minimise it. Defaults to false.
* autoToggleDisplay: Sets whether the minimap should hide automatically, if the parent map bounds does not fit within the minimap bounds. Especially useful when ’zoomLevelFixed’
is set.
* ``minimized``: Sets whether the minimap should start in a minimized position.
* ``aimingRectOptions``: Sets the style of the aiming rectangle by passing in a Path.Options (http://leafletjs.com/reference-1.3.1.html#path-options) object. (Clickable
will always be overridden and set to false.)
* shadowRectOptions
Sets the style of the aiming shadow rectangle by passing in a Path.Options (http://leafletjs.com/reference-1.3.1.html#path-option) object. (Clickable
will always be overridden and set to false.)
* ``strings``: Overrides the default strings allowing for translation.
* ``tiles``: URL for tiles or one of the pre-defined providers.
* mapOptions Sets Leaflet options for the MiniMap map. It does not override the MiniMap default map options but extends them.


### Examples

```{r}
leaf <- leaflet() %>%
addTiles() %>%
addMiniMap()

leaf
```

#### See Also
providers
