---
title: "leaflet: ``setView`` Methods to manipulate the map widget"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
```

## Description

A series of methods to manipulate the map.
### Usage

<pre><code>
setView(map, lng, lat, zoom, options = list())
flyTo(map, lng, lat, zoom, options = list())
fitBounds(map, lng1, lat1, lng2, lat2, options = list())
flyToBounds(map, lng1, lat1, lng2, lat2, options = list())
setMaxBounds(map, lng1, lat1, lng2, lat2)
clearBounds(map)
</code></pre>

### Arguments

* ``map``: a map widget object created from leaflet()
* ``lng``: The longitude of the map center
* ``lat``: The latitude of the map center
* ``zoom``: the zoom level
* ``options``: a list of zoom/pan options (see http://leafletjs.com/reference-1.3.1.html#zoom/pan-options)
* ``lng1, lat1, lng2, lat2``: the coordinates of the map bounds



### Functions

*  setView: Set the view of the map (center and zoom level)
*  flyTo: Flys to a given location/zoom-level using smooth pan-zoom.
*  fitBounds: Set the bounds of a map
*  flyToBounds: Flys to given bound using smooth pan/zoom.
*  setMaxBounds: Restricts the map view to the given bounds
*  clearBounds: Clear the bounds of a map, and the bounds will be automatically determined from latitudes and longitudes of the map elements if available (otherwise the full world view is used)

### References

http://leafletjs.com/reference-1.3.1.html#map-methods-for-modifying-map-state

### Examples

```{r}
m <- leaflet() %>% addTiles() %>% setView(-71.0382679, 42.3489054, zoom = 18)
m # the RStudio 'headquarter'
m %>% fitBounds(-72, 40, -70, 43)
m %>% clearBounds() # world view
```
#### Value
The modified map widget.
